---
title: "Lab 1: Explore & Get Started in R"
subtitle: "Lab 1 - PSY 150"
author: "ADD_YOUR_NAME_HERE"
date: "`r Sys.Date()`"
output: 
    html_document:
      theme: sketchy
---

```{r setup, echo=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

------------------------------------------------------------------------

Step 1: Load packages
```{r}

library(tidyverse)
library(here)
library(janitor)
library(ggthemes)
library(viridis)
library(ThemePark)

```

------------------------------------------------------------------------

### Keyboard shortcuts (Use them!)

> WINDOWS 

- `Run code` (at cursor): CONTROL + ENTER 
- `Pipe operator` (`%>%`): COMMAND + SHIFT + M
- `Add code chunk`: COMMAND + OPTION + I
- `Assignment operator` (`<-`): OPTION + MINUS(-) 
- `Search`: COMMAND + F

------------------------------------------------------------------------

### How this lab will work!

- Work from top to bottom. 
- Run a *code chunk* (grey box) by clicking the green triangle (upper right corner) OR by placing your cursor in the chunk and pressing **Ctrl + Shift + Enter** (Windows)


### Run the next `code chunk`

Rectangles beginning and ending with three backticks contain R code. Running a chunk executes its lines in order and displays the results.

```{r}

cat("Hello, World!\n")

```


The fundamental building block of R code is an **expression**. 

- A code chunk can contain multiple expressions 
- R runs them from top to bottom
- You can also include comments in a code block using the `#` symbol 

```{r}
# This is a comment, not R code!

cat("First this line is printed,\n")
cat("then the whole 🌏,\n")
cat("and then this one.\n")
```

### Adding a new code chunk

You can add a new R chunk by clicking the green +C button above OR by using the keyboard shortcut **Ctrl + Alt + I** (Windows).

Add a new R code chunk below this paragraph. Use `cat()` to print:

``` text
A whole new chunk! ♪ 🌏 ♪
```

Run your new chunk to verify that it works.



### Catching Errors in R

- R is a language with precise rules. 
- Small mistakes may produce an error message. 
- Errors are normal: experienced programmers make them all the time. 
- The goal is to read the message, find the problem, and try again.

The next chunk is intentionally incomplete. Run the chunk, and read the error message. Then add the missing right parenthesis and run it again.

```{r}
# Fix the coding error 

cat("This line is missing something."
    
```

------------------------------------------------------------------------

### Read in the class survey data

------------------------------------------------------------------------

```{r}

raw_data <- *** 

  
```

```{r}

 class_data <- raw_data %>% 
  mutate(across(c(fav_number, fav_season, pets, dominant_hand, risk_pref, study_sound),
                as.factor))

```

------------------------------------------------------------------------

### Exploring classroom height

Survey question: *What is your height (inches)?* 

------------------------------------------------------------------------

Take a look at the values in the variable `height` within the `data.frame` named `class_data`
```{r}



```


------------------------------------------------------------------------

Next let's look at `height` as a frequency table
```{r}



```

------------------------------------------------------------------------

Make a fast histogram using the function `hist()`
```{r}



```

------------------------------------------------------------------------

#### Make a histogram using `ggplot` for variable `height`

```{r}



```

------------------------------------------------------------------------

### R is just a fancy calculator - measures of central tendency

1.  Calculate the mean ("the average")
2.  Find the median ("the middle")
3.  Find the mode ("the most common")

------------------------------------------------------------------------

#### 1. Calculate the average (`mean`)

```{r}



```

There is also a function for that:
```{r}



```

------------------------------------------------------------------------

#### 2. Finding the `median`

```{r}
# Step 1: Order values from lowest to highest


# Step 2: Find the middle value(s)
```

**Response**:_________ADD_YOUR_RESPONSE___________

------------------------------------------------------------------------

There is also a function for that:
```{r}



```

------------------------------------------------------------------------

#### 3. Finding the `mode` ("most common value")

**NOTE:** There can be multiple modes (if there is more than 1 highest frequency)

```{r}

# Step 1: Make a frequency table


# Step 2: Find the highest frequency 
```

**Response**:_________ADD_YOUR_RESPONSE___________

------------------------------------------------------------------------

#### Hmmm... there is no function for that.

So let's write a simple function ourselves!

```{r}

# Step 1: Write your own function



# Step 2: Plug the input into your new function



```

------------------------------------------------------------------------

### Favorite number

Survey question: *What is your favorite number between 1 and 10?* 

------------------------------------------------------------------------

```{r}



```

::: callout
`Question`: What measure of central tendency would be best for this type of variable?
:::

**Response**:_________ADD_YOUR_RESPONSE___________

------------------------------------------------------------------------

### 🍂 Favorite season

Survey question: *What is your favorite season?* 

------------------------------------------------------------------------

```{r}

class_data %>%
  ggplot(aes(x = , fill = )) +
  geom_bar() +
  labs(title = "What is your favorite season?",
  x = "Favorite Season",
  y = "Count") +
  guides(fill = "none")

```

::: callout
`Question`: When do we use `geom_bar()` versus `geom_histogram()`?
:::

**Response**:_________ADD_YOUR_RESPONSE___________

------------------------------------------------------------------------

### ⏰ Wake up time

Survey question: *What is your preferred wake up time?* 

------------------------------------------------------------------------

```{r}

class_data %>%
  # mutate(wake_up_time = 
  #        factor(wake_up_time,
  #        levels = c("6 am", "7 am", "8 am", "9 am", "10 am",
  #                    "11 am", "12 pm", "1 pm (or later)"))) %>%
  ggplot(aes(x = , fill = )) +
  geom_bar() +
  labs( title = "Preferred wake up time?",
    x = "Wake up time",
    y = "Student Count") +
  guides(fill = "none")

```

------------------------------------------------------------------------

### ✋ Dominant hand

Survey question: *Which is your dominant hand?* 

------------------------------------------------------------------------

```{r}

class_data %>% 
ggplot(aes(x = , fill = )) +
  geom_bar() +
  labs(
    title = "Which is your dominant hand?",
    x = "",
    y = "Count") +
  theme(legend.position = "none")

```

::: callout
`Question`: What can we do with a variable with no (or little) variation?
:::

**Response**:_________ADD_YOUR_RESPONSE___________

------------------------------------------------------------------------

### 🐶 Pets

Survey question: *Do you have any pets?* 

------------------------------------------------------------------------

```{r}
class_data %>% 
ggplot(aes(x = , fill = )) +
  geom_bar() +
  labs(
    title = "Do you have any pets?",
    x = "",
    y = "Student Count") +
  theme(legend.position = "none")
```

------------------------------------------------------------------------

### 🎲 Risk/reward preference

Survey question: *Which would you choose?* 

a. *A sure $10*
b. *A 50% chance to win $24*

------------------------------------------------------------------------

```{r}
class_data %>% 
  #mutate(risk_pref = fct_rev(risk_pref)) %>% 
ggplot(aes(x = , fill = )) +
  geom_bar() +
  labs(
    title = "Risk preference",
    subtitle = "Which would you choose?",
    x = "",
    y = "Student Count") +
  theme(legend.position = "none")
```

------------------------------------------------------------------------

### 🎶 Study sound

Survey question: *Do you prefer background sound while studying?* 

------------------------------------------------------------------------

```{r}

class_data %>% 
ggplot(aes(x = , fill = )) +
  geom_bar() +
  labs(
    subtitle = "Do you prefer background sound while studying?",
    x = "",
    y = "Student Count"
  ) +
  theme(legend.position = "none")

```


------------------------------------------------------------------------

