---
title: "Code-along â†’ Explore & Get Started in R"
subtitle: "Lab: PSY 150"
author: "ADD_YOUR_NAME_HERE"
date: "`r Sys.Date()`"
output: 
    html_document:
      theme: sketchy
---

```{r setup, echo=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

------------------------------------------------------------------------

### ðŸ“¦ Packages: Install once... load libraries every time!

-   If a `yellow banner` appears at top of screen, asking you to install packages click `YES`
-   Packages include functions other R users have made for us to make our lives easier!

------------------------------------------------------------------------

Load packages
```{r}

library(tidyverse)
library(here)
library(janitor)
library(ggthemes)
library(ggridges) 
library(viridis)
library(ThemePark)

```

------------------------------------------------------------------------

### Keyboard shortcuts (Use them!)

> MAC / *WINDOWS*

- `Run code` (at cursor): COMMAND/*CONTROL* + RETURN/*ENTER*
- `Pipe operator` (`|>`): COMMAND/*CONTROL* + SHIFT + M
- `Add code chunk`: COMMAND/*CONTROL* + OPTION/*ALT* + I
- `Assignment operator` (`<-`): OPTION/*ALT* + MINUS("-") 
- `Search`: COMMAND/*CONTROL* + F

------------------------------------------------------------------------

### Read in the class survey data

------------------------------------------------------------------------

```{r}

class_data <-  
  mutate(across(c(fav_number, fav_season, pets, dominant_hand, risk_pref, study_sound),
                as.factor))
  
```

------------------------------------------------------------------------

### Exploring classroom height

Survey question: *What is your height (inches)?* 

------------------------------------------------------------------------

Take a look at the values in the variable `height` within the `data.frame` named `class_data`
```{r}



```

::: callout
`NOTE`: This list of number is called a `vector`
:::

------------------------------------------------------------------------

Next let's look at `height` as a frequency table
```{r}



```

------------------------------------------------------------------------

Make a fast histogram using the function `hist()`
```{r}



```

------------------------------------------------------------------------

#### Make a histogram using `ggplot` for variable `height`

```{r}



```

------------------------------------------------------------------------

### R is just a fancy calculator - measures of central tendency

1.  Calculate the mean ("the average")
2.  Find the median ("the middle")
3.  Find the mode ("the most common")

------------------------------------------------------------------------

#### 1. Calculate the average (`mean`)

```{r}



```

There is also a function for that:
```{r}



```

------------------------------------------------------------------------

#### 2. Finding the `median`

```{r}
# Step 1: Order values from lowest to highest


# Step 2: Find the middle value(s)
```

**Response**:_________ADD_YOUR_RESPONSE___________

------------------------------------------------------------------------

There is also a function for that:
```{r}



```

------------------------------------------------------------------------

#### 3. Finding the `mode` ("most common value")

**NOTE:** There can be multiple modes (if there is more than 1 highest frequency)

```{r}

# Step 1: Make a frequency table


# Step 2: Find the highest frequency 
```

**Response**:_________ADD_YOUR_RESPONSE___________

------------------------------------------------------------------------

#### Hmmm... there is no function for that.

So let's write a simple function ourselves!

```{r}

# Step 1: Write your own function



# Step 2: Plug the input into your new function



```

------------------------------------------------------------------------

### Favorite number

Survey question: *What is your favorite number between 1 and 10?* 

------------------------------------------------------------------------

```{r}



```

::: callout
`Question`: What measure of central tendency would be best for this type of variable?
:::

**Response**:_________ADD_YOUR_RESPONSE___________

------------------------------------------------------------------------

### ðŸ‚ Favorite season

Survey question: *What is your favorite season?* 

------------------------------------------------------------------------

```{r}

class_data |>
  ggplot(aes(x = , fill = )) +
  geom_bar() +
  labs(title = "What is your favorite season?",
  x = "Favorite Season",
  y = "Count") +
  guides(fill = "none")

```

::: callout
`Question`: When do we use `geom_bar()` versus `geom_histogram()`?
:::

**Response**:_________ADD_YOUR_RESPONSE___________

------------------------------------------------------------------------

### â° Wake up time

Survey question: *What is your preferred wake up time?* 

------------------------------------------------------------------------

```{r}

class_data |>
  # mutate(wake_up_time = 
  #        factor(wake_up_time,
  #        levels = c("6 am", "7 am", "8 am", "9 am", "10 am", "11 am", "12 pm", "1 pm (or later)"),
  #        ordered = TRUE)) |>
  ggplot(aes(x = , fill = )) +
  geom_bar() +
  labs( title = "Preferred wake up time?",
    x = "Wake up time",
    y = "Student Count") +
  guides(fill = "none")

```

------------------------------------------------------------------------

### âœ‹ Dominant hand

Survey question: *Which is your dominant hand?* 

------------------------------------------------------------------------

```{r}

class_data |> 
ggplot(aes(x = , fill = )) +
  geom_bar() +
  labs(
    title = "Which is your dominant hand?",
    x = "",
    y = "Count") +
  theme(legend.position = "none")

```

::: callout
`Question`: What can we do with a variable with no (or little) variation?
:::

**Response**:_________ADD_YOUR_RESPONSE___________

------------------------------------------------------------------------

### ðŸ¶ Pets

Survey question: *Do you have any pets?* 

------------------------------------------------------------------------

```{r}
class_data |> 
ggplot(aes(x = , fill = )) +
  geom_bar() +
  labs(
    title = "Do you have any pets?",
    x = "",
    y = "Student Count") +
  theme(legend.position = "none")
```

------------------------------------------------------------------------

### ðŸŽ² Risk/reward preference

Survey question: *Which would you choose?* 

a. *A sure $10*
b. *A 50% chance to win $24*

------------------------------------------------------------------------

```{r}
class_data |> 
  #mutate(risk_pref = fct_rev(risk_pref)) |> 
ggplot(aes(x = , fill = )) +
  geom_bar() +
  labs(
    title = "Risk preference",
    subtitle = "Which would you choose?",
    x = "",
    y = "Student Count") +
  theme(legend.position = "none")
```

------------------------------------------------------------------------

### ðŸŽ¶ Study sound

Survey question: *Do you prefer background sound while studying?* 

------------------------------------------------------------------------

```{r}

class_data |> 
ggplot(aes(x = , fill = )) +
  geom_bar() +
  labs(
    subtitle = "Do you prefer background sound while studying?",
    x = "",
    y = "Student Count"
  ) +
  theme(legend.position = "none")

```


------------------------------------------------------------------------

